Mermory-efficient FO transform using ``encrypt-then-MAC''

### 2024-05-30 lattices/compact-fo-using-mac/manuscript/main.tex
Checkpoint for Friday meeting

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\Norm}[1]{\Vert {#1} \Vert}
\newcommand{\norm}[1]{\vert {#1} \vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}
\newcommand{\keygen}{\operatorname{KeyGen}}
\newcommand{\pk}{\operatorname{pk}}
\newcommand{\sk}{\operatorname{sk}}
\newcommand{\pco}{\operatorname{PCO}}
\newcommand{\cvo}{\operatorname{CVO}}
\newcommand{\coin}{\operatorname{COIN}}
\newcommand{\encap}{\operatorname{Encap}}
\newcommand{\decap}{\operatorname{Decap}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newlength{\wdth}
\newcommand{\strike}[1]{\settowidth{\wdth}{#1}\rlap{\rule[.5ex]{\wdth}{.4pt}}#1}

@@

\section{Preliminaries}
\subsection{Security definitions}
The security of public-key encryption schemes (PKE) and message authentication code (MAC) is defined using an number of adversarial games, where each game is defined by the adversary's goal and objectives.

For PKE, we define the $\operatorname{OW-ATK}$ and $\operatorname{IND-ATK}$ security game as follows:

\begin{multicols}{2}
    \begin{algorithm}[H]
        \caption{OW-ATK security game}
        \KwIn{
            Security parameter $\lambda$, adversary $\mathcal{A}$
        }
        \SetAlgoLined
        $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
        $m^\ast \leftsample \mathcal{M}$\;
        $c^\ast \leftsample E(\pk, m^\ast)$\;
        $\hat{m} \leftarrow \mathcal{A}(1^\lambda, \pk, c^\ast, \mathcal{O}_{\operatorname{ATK}})$\;
        \Return{
            $\llbrack m = \hat{m} \rrbrack$
        }
    \end{algorithm}

    \begin{algorithm}[H]
        \caption{IND-ATK security game}
        \KwIn{
            Security param $\lambda$, adversary $\mathcal{A}$
        }
        \SetAlgoLined
        $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
        $(m_0, m_1) \leftsample \mathcal{A}(1^\lambda, \pk, \mathcal{O}_{\operatorname{ATK}})$\;
        $b \leftsample \{0, 1\}$\;
        $c^\ast \leftarrow E(\pk, m_b)$\;
        $\hat{b} \leftarrow \mathcal{A}(1^\lambda, \pk, c^\ast, \mathcal{O}_{\operatorname{ATK}})$\;
        \Return{
            $\llbrack \hat{b} = b \rrbrack$
        }
    \end{algorithm}
\end{multicols}

@@

\subsection{$\delta$-correctness}
We define the correctness of a PKE using an adversarial game in which an adversary tries to produce plaintext that triggers decryption failures.

\begin{algorithm}[H]
    \caption{Correctness game $\operatorname{CORS}$}
    \SetAlgoLined
    $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
    $m \leftarrow \mathcal{A}(1^\lambda, \pk)$\;
    \Return{
        $\llbrack D(\sk, E(\pk, m)) \neq m \rrbrack$\;
    }
\end{algorithm}

\begin{definition}
    A public-key encryption scheme $\mathcal{E}=(E, D)$ is $\delta$-correct if for any efficient adversary $\mathcal{A}$, the probability of winning the correctness game is bounded by $\delta$:

    \begin{equation*}
        P[\operatorname{CORS}(\mathcal{A}) \rightarrow 1] \leq \delta
    \end{equation*}
\end{definition}

>>>>

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\Norm}[1]{\Vert {#1} \Vert}
\newcommand{\norm}[1]{\vert {#1} \vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}
\newcommand{\keygen}{\operatorname{KeyGen}}
\newcommand{\pk}{\operatorname{pk}}
\newcommand{\sk}{\operatorname{sk}}
\newcommand{\pco}{\operatorname{PCO}}
\newcommand{\cvo}{\operatorname{CVO}}
\newcommand{\coin}{\operatorname{COIN}}
\newcommand{\encap}{\operatorname{Encap}}
\newcommand{\decap}{\operatorname{Decap}}
\newcommand{\pke}{\operatorname{PKE}}
\newcommand{\mac}{\operatorname{MAC}}
\newcommand{\cors}{\operatorname{CORS}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newlength{\wdth}
\newcommand{\strike}[1]{\settowidth{\wdth}{#1}\rlap{\rule[.5ex]{\wdth}{.4pt}}#1}

@@

\section{Preliminaries}
\subsection{Public-key encryption}
A public key encryption scheme $\pke$ consists of three algorithms $(\keygen, E, D)$ defined over the collection of key space, message space, and ciphertext space $(\mathcal{K}, \mathcal{E}, \mathcal{D})$. The $\keygen$ routines takes the security parameter $1^\lambda$ and outputs a keypair $(\pk, \sk)$. The encryption routine $E$ takes a public key $\pk$ and a plaintext message $m \in \mathcal{M}$, and outputs a ciphertext $c \in \mathcal{C}$. Where the encryption routine is probabilistic, we denote the source of randomness by a coin sampled from the coin space $r \in \mathcal{R}$ such that $c \leftsample E(\pk, m; r)$. When $m, r$ is both specified, $E$ must be deterministic and injective. The decryption routine $D(\sk, c)$ takes a secret key and a ciphertext, and outputs the corresponding plaintext. Decryption routine is always deterministic. Where appropriate, the decryption routine could also output the rejection symbol $\bot$ to indicate that the ciphertext is invalid.

Correctness is commonly required of all $\pke$'s, which means that for all possible key pairs $(\pk, \sk)$ and messages $m \in \mathcal{M}$, $D(\sk, E(\sk, m)) = m$. However, with lattice-based schemes (such as CRYSTALS-Kyber), for randomly sampled keypairs and messages, there is a non-zero probability that $D(\sk, E(\pk, m)) \neq m$. Furthermore, decryption failures could be an attack vector through which an adversary learns partial information about the keys or ciphertext\cite{bos2018crystals}. To account for this possibility, we define a correctness game, then model the probability of decryption failure on this game.

In the correctness game $\cors$, an adversary is given the security parameter and some public key and outputs a plaintext message and a coin. The adversary wins the $\cors$ game if and only if $m$ triggers a decryption failure for the specified keypair. Note that this game can be trivially adapted for deterministic encryption scheme, so we will skip the discussion here.

\begin{algorithm}[H]
    \caption{The correctness game $\cors$}
    \KwIn{
        The correctness adversary $\mathcal{A}$
    }
    $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
    $(m, r) \leftsample \mathcal{A}(1^\lambda, \pk)$\;
    \Return{
        $\llbrack D(\sk, E(\pk, m; r)) \neq m \rrbrack$
    }
\end{algorithm}

\begin{definition}[$\delta$-correctness]\label{def:delta-correctness}
    A public-key encryption scheme $\pke$ is $\delta$-correct if no (possibly unbounded) adversary can win the $\cors$ game with probability greater than $\delta$.
\end{definition}

For some concrete instantiation, the value of $\delta$ for Kyber\cite{avanzi2019crystals} is shown in table \ref{table:kyber-delta-vals}:

\begin{table}
    \centering
    \begin{tabular}{cc}
        Security level & $\delta$ \\
        Kyber512 & $2^{-139}$ \\
        Kyber768 & $2^{-164}$ \\
        Kyber1024 & $2^{-174}$ \\
    \end{tabular}
    \caption{Concrete $\delta$ value for CRYSTAL-Kyber}
    \label{table:kyber-delta-vals}
\end{table}

The security of a $\pke$ is defined using an number of adversarial games, where each game is defined by the adversary's goal and objectives. We define the $\operatorname{OW-ATK}$ and $\operatorname{IND-ATK}$ security game in algorithm \ref{alg:ow-atk-game} and algorithm \ref{alg:ind-atk-game}:

\begin{multicols}{2}
    \begin{algorithm}[H]\label{alg:ow-atk-game}
        \caption{OW-ATK security game}
        \KwIn{
            Security parameter $\lambda$, adversary $\mathcal{A}$
        }
        \SetAlgoLined
        $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
        $m^\ast \leftsample \mathcal{M}$\;
        $c^\ast \leftsample E(\pk, m^\ast)$\;
        $\hat{m} \leftarrow \mathcal{A}(1^\lambda, \pk, c^\ast, \mathcal{O}_{\operatorname{ATK}})$\;
        \Return{
            $\llbrack m = \hat{m} \rrbrack$
        }
    \end{algorithm}

    \begin{algorithm}[H]\label{alg:ind-atk-game}
        \caption{IND-ATK security game}
        \KwIn{
            Security param $\lambda$, adversary $\mathcal{A}$
        }
        \SetAlgoLined
        $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
        $(m_0, m_1) \leftsample \mathcal{A}(1^\lambda, \pk, \mathcal{O}_{\operatorname{ATK}})$\;
        $b \leftsample \{0, 1\}$\;
        $c^\ast \leftarrow E(\pk, m_b)$\;
        $\hat{b} \leftarrow \mathcal{A}(1^\lambda, \pk, c^\ast, \mathcal{O}_{\operatorname{ATK}})$\;
        \Return{
            $\llbrack \hat{b} = b \rrbrack$
        }
    \end{algorithm}
\end{multicols}

@@

\subsection{Message authentication code}
A message authentication code $\mac$ consists of a pair of routines $(S, V)$ defined over a collection $(\mathcal{K}, \mathcal{M}, \mathcal{T})$ of key space, message, space, and tag space. The signing routine $S(k, m)$ takes the secret key $k$ and a message $m$, and outputs a tag $t$. The verfiying routine $V(k, m, t)$ takes the triplet of secret key $k$, message $m$, and tag $t$, and outputs 1 if the message-tag pair is valid for the given secret key or 0 otherwise.

One common notion of security for $\mac$ is ``existential forgery under chosen message attack'' (EF-CMA), which is defined using an adversarial game. In the EF-CMA game, an adversary is given access to a signing oracle $\mathcal{O}^S$ and tries to forge valid message-tag pairs that have not be recorded in the signing oracle before.

\begin{multicols}{2}
    \begin{algorithm}[H]\label{alg:euf-cma-game}
        \caption{
            EF-CMA game
        }
        $k^\ast \leftsample \mathcal{K}$\;
        $(m, t) \leftarrow \mathcal{A}^{\mathcal{O}^S}()$\;
        \Return{
            $\llbrack (m, t) \not\in \mathcal{O}^{S} \rrbrack$
            and $\llbrack V(k^\ast, m, t) \rrbrack$
        }
    \end{algorithm}

    \begin{algorithm}[H]\label{alg:mac-signing-oracle}
        \caption{The signing oracle $\mathcal{O}^S$}
        \KwIn{
            $m \in \mathcal{M}$
        }
        $t \leftarrow S(k^\ast, m)$\;
        \If{$(m, t) \not\in \mathcal{O}^S$}{
            Append $(m, t)$ to $\mathcal{O}^S$
        }
        \Return{$t$}
    \end{algorithm}
\end{multicols}

\begin{definition}[Existential unforgeability]\label{def:euf-cma}
    A $\mac$ is existentially unforgeable under chosen message attack if for all efficient adversary, the probability of winning the EF-CMA game is negligible with respect to the security parameter.
\end{definition}
